---
title: "roxygen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{roxygen}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
show_file <- function(path, lang) {
  writeLines(c(paste0("```", lang), readLines(path), "```"))
}
porcelain_file <- function(path) {
  system.file(path, package = "porcelain", mustWork = TRUE)
}
```

The approach described in `vignette("porcelain")` allows maximum flexibility, but `porcelain` also allows a less programmatic, more declarative approach via custom roxygen tags.  This approach returns to the original expressiveness of `plumber`, but with the type-checking and testability of `porcelain`.

## Prerequisites

You must first update your `DESCRIPTION` file to to add the line

```
Roxygen: list(roclets = c("rd", "namespace", "porcelain::porcelain_roclet"))
```

or if you are using roxygen-with-markdown

```
Roxygen: list(markdown = TRUE, roclets = c("rd", "namespace", "porcelain::porcelain_roclet"))
```

The important part is the `porcelain::porcelain_roclet` entry in "roclets" which will activate the `@porcelain` tag that we will use.

## Declaring a simple endpoint

Consider our simple `add` function from `vignette("porcelain")`, which takes numeric query arguments and returns a number.  We can write

```r
#' @porcelain
#'   GET / => json("numeric")
#'   query a :: numeric
#'   query b :: numeric
add <- function(a, b) {
  jsonlite::unbox(a + b)
}

api <- function(validate = FALSE) {
  api <- porcelain::porcelain$new(validate = validate)
  api$handle_package()
  api
}
```

Then running `devtools::document()` (or similar) will generate the file `R/porcelain.R` with a hook function that contains code that will be included when running the `handle_package` method.

## The basic syntax

The basic tag parts required are

```r
#' @porcelain <METHOD> <path> => <returning>
```

This string can be split across many lines and additional whitespace is ignored, so for the simple example above, we could equivalently write

```
#' @porcelain
#'   GET /add/<a:int>/<b:int> =>
#'     json("numeric")
```

The arguments are all passed to `porcelain::porcelain_endpoint`; the method and path are passed directly as `method` and `path`. The `=>` symbol exists to read "returning" and help mark the end of the path.

The "returning" argument is transformed before being passed through. In the example above `json("numeric")` became `porcelain::porcelain_returning_json("numeric")`

We translate with the rule:

* `json`: `porcelain::porcelain_returning_json`
* `binary`: `porcelain::porcelain_returning_binary`
* `generic` `porcelain::porcelain_returning`

If if no arguments are provided we perform a call with no arguments (e.g., `GET /path => json` would use `porcelain::porcelain_returning_json()`). If arguments are given, then we will attempt to auto-quote these.

## Adding inputs

The above function used query parameters to provide inputs to the target function; you can specify query, body and bound state this way (path parameters remain in the path, as before).

All input parameters have the format

```
#' <location> <name> :: <description>
```

where

* `<location>` is one of `query`, `body` or `state`
* `<name>` is the name of the argument in the target function that you will send the input to
* `<description>` is one or more arguments that describes the input further

 The interface here is a little different to the underlying porcelain functions (`porcelain::porcelain_input_query`, `porcelain::porcelain_input_body` and `porcelain::porcelain_state`) in that every input parameter is given individually even if you might treat these together in the functions.  For example, for two input queries we used two lines.

**For query inputs**, the description will be one of the valid types for `porcelain::porcelain_input_query`; `logical`, `integer`, `numeric` or `string`, for example an endpoint accepting two query inputs:

```r
#' @porcelain
#'   POST /path => json(OutputSchema)
#'   query a :: numeric
#'   query b :: integer
f <- function(a, b) {
  # implementation
}
```

**For a json body** it is likely that you will want to add a schema

```r
#' @porcelain
#'   POST /path => json(OutputSchema)
#'   query a :: numeric
#'   query b :: integer
#'   body data :: json(InputSchema)
f <- function(a, b, data) {
  ...
}
```

**For a binary body** you can optionally specify the incoming mime type, for example

```
#'   body data :: binary(application/zip)
```

would refer to a zip input.

## Binding state

Finally consider binding state into the API

## Testing

Because the code is generated into the porcelain.R file you need to use `porcelain::porcelain_package_endpoint` in order to extract the raw endpoint object to take advantage of porcelain's test helper.

```r
endpoint <- porcelain::porcelain_package_endpoint("mypkg", "GET", "/path")
endpoint$run()
```

## An example

We include a very small complete example

```{r, include = FALSE}
tmp <- tempfile()
dir.create(tmp, FALSE, TRUE)
file.copy(porcelain_file("examples/add2"), tmp, recursive = TRUE)
path <- file.path(tmp, "add2")
```


```{r, echo = FALSE, comment = NA}
withr::with_dir(dirname(path), fs::dir_tree("add2"))
```

As for the simple example in `vignette("porcelain")` we have written the api into a single file `R/api.R` but this could be split over as many files as you prefer

```{r, echo = FALSE, results = "asis"}
show_file(file.path(path, "R/api.R"), "r")
```

Our `DESCRIPTION` file includes the roxygen2 setup

```{r, echo = FALSE, results = "asis"}
show_file(file.path(path, "DESCRIPTION"), "")
```

We include a small json schema `inst/schema/numeric.json`:

```{r, echo = FALSE, results = "asis"}
show_file(file.path(path, "inst/schema/numeric.json"), "json")
```

Running `roxygen2::roxygenize(path)` on the file will build the interface

```{r}
roxygen2::roxygenize(path)
```

The contents now include `R/porcelain.R` (and `man`, created by roxygen2 but empty)

```{r, echo = FALSE, comment = NA}
withr::with_dir(dirname(path), fs::dir_tree("add2"))
```

The `R/porcelain.R` file contains automatically-generated endpoint definitions

```{r, echo = FALSE, results = "asis"}
show_file(file.path(path, "R/porcelain.R"), "r")
```
