---
title: "roxygen"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{roxygen}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
show_file <- function(path, lang) {
  writeLines(c(paste0("```", lang), readLines(path), "```"))
}
porcelain_file <- function(path) {
  system.file(path, package = "porcelain", mustWork = TRUE)
}
```

The approach described in `vignette("porcelain")` allows maximum flexibility, but `porcelain` also allows a less programmatic, more declarative approach via custom roxygen tags.  This approach returns to the original expressiveness of `plumber`, but with the type-checking and testability of `porcelain`.

## Prerequisites

You must first update your `DESCRIPTION` file to to add the line

```
Roxygen: list(roclets = c("rd", "namespace", "porcelain::porcelain_roclet"))
```

or if you are using roxygen-with-markdown

```
Roxygen: list(markdown = TRUE, roclets = c("rd", "namespace", "porcelain::porcelain_roclet"))
```

The important part is the `porcelain::porcelain_roclet` entry in "roclets" which will activate the `@porcelain` tag that we will use.

## Declaring a simple endpoint

Consider our simple `add` function from `vignette("porcelain")`, which takes numeric query arguments and returns a number.  We can write

```r
#' @porcelain
#'   GET / => json("numeric")
#'   query a :: numeric
#'   query b :: numeric
add <- function(a, b) {
  jsonlite::unbox(a + b)
}

api <- function(validate = FALSE) {
  api <- porcelain::porcelain$new(validate = validate)
  api$handle_package()
  api
}
        ```

Then running `devtools::document()` (or similar) will generate the file `R/porcelain.R` with a hook function that contains code that will be included when running the `handle_package` method.

## The basic syntax

The basic tag parts required are

```r
#' @porcelain <METHOD> <path> => <returning>
```

This string can be split across many lines and additional whitespace is ignored, so for the simple example above, we could equivalently write

```
#' @porcelain
#'   GET /add/<a:int>/<b:int> =>
#'     json("numeric")
```

The arguments are all passed to `porcelain::porcelain_endpoint`; the method and path are passed directly as `method` and `path`. The `=>` symbol exists to read "returning" and help mark the end of the path.

The "returning" argument is transformed before being passed through. In the example above `json("numeric")` became `porcelain::porcelain_returning_json("numeric")`

We translate with the rule:

* `json`: `porcelain::porcelain_returning_json`
* `binary`: `porcelain::porcelain_returning_binary`
* `generic` `porcelain::porcelain_returning`

If if no arguments are provided we perform a call with no arguments (e.g., `GET /path => json` would use `porcelain::porcelain_returning_json()`). If arguments are given, then we will attempt to auto-quote these.

## Adding inputs

The above function used query parameters to provide inputs to the target function; you can specify query, body and bound state this way (path parameters remain in the path, as before).

All input parameters have the format

```
#' <location> <name> :: <description>
```

where

* `<location>` is one of `query`, `body` or `state`
* `<name>` is the name of the argument in the target function that you will send the input to
* `<description>` is one or more arguments that describes the input further

 The interface here is a little different to the underlying porcelain functions (`porcelain::porcelain_input_query`, `porcelain::porcelain_input_body` and `porcelain::porcelain_state`) in that every input parameter is given individually even if you might treat these together in the functions.  For example, for two input queries we used two lines.

**For query inputs**, the description will be one of the valid types for `porcelain::porcelain_input_query`; `logical`, `integer`, `numeric` or `string`, for example an endpoint accepting two query inputs:

```r
#' @porcelain
#'   POST /path => json(OutputSchema)
#'   query a :: numeric
#'   query b :: integer
f <- function(a, b) {
  # implementation
}
```

**For a json body** it is likely that you will want to add a schema

```r
#' @porcelain
#'   POST /path => json(OutputSchema)
#'   query a :: numeric
#'   query b :: integer
#'   body data :: json(InputSchema)
f <- function(a, b, data) {
  ...
}
```

**For a binary body** you can optionally specify the incoming mime type, for example

```
#'   body data :: binary(application/zip)
```

would refer to a zip input.

No special markup is required for path parameters, include these using plumber's syntax

```r
#' @porcelain
#'   POST /path/<a>/<b:int> => json(OutputSchema)
f <- function(a, b) {
  ...
}
```

## Binding state

Finally consider binding state into the API.  We need to do this where we have some mutable state in the API that endpoints (typically `POST` or `DELETE`) will modify.  Examples include database connections or queues.

Suppose that you have an endpoint that will count the number of times that it has been accessed, returning that number. We might use a counter like this:

```{r}
counter <- R6::R6Class(
  "counter",
  private = list(n = 0),
  public = list(
    value = function() {
      private$n
    },
    increase = function() {
      private$n <- private$n + 1
      private$n
    }))
```

Outside of porcelain we can use this like so:

```{r}
obj <- counter$new()
obj$increase()
obj$increase()
obj$value()
```

We can write a set of endpoints that that share a `counter` object like this, and add roxygen comments to configure it:

```r
#' @porcelain
#'   GET /counter/value => json(number)
#'   state obj :: counter
increase <- function(obj) {
  jsonlite::unbox(obj$value())
}

#' @porcelain
#'   POST /counter/increase => json(number)
#'   state obj :: counter
value <- function(obj) {
  jsonlite::unbox(obj$increase())
}
```

Here we have a pair of endpoints; the first `GET /counter/value` will return the value of the counter and `POST /counter/increase` will increase its value by one and return that. We could have other methods like `POST /counter/reset` to reset the counter, but the key thing is that all endpoints must *share the same counter* and to do that we must pass it to the api when we create it.

Here we say that doing `POST /counter` will return a number. The endpoint takes a `counter` object as above, but that won't come from the HTTP request - it will be bound into the API, so it might be shared.  By adding the roxygen comment

```
#'   state obj :: counter
```

we set this up.  We need to pass the state through when creating the API:

```
api <- function(validate = FALSE) {
  state <- list(counter = counter$new())
  api <- porcelain::porcelain$new(validate = validate)
  api$handle_package(state)
  api
}
```

Here our api creation function creates a new zeroed counter (you could accept one as an argument of course), puts that into a list with name `counter`, corresponding to the rhs of the roxygen comment, and passes that through to `handle_package`.

## Testing

Because the code is generated into the porcelain.R file you need to use `porcelain::porcelain_package_endpoint` in order to extract the raw endpoint object to take advantage of porcelain's test helper.

```r
endpoint <- porcelain::porcelain_package_endpoint("mypkg", "GET", "/path")
endpoint$run()
```

## An example

We include a very small complete example

```{r, include = FALSE}
tmp <- tempfile()
dir.create(tmp, FALSE, TRUE)
file.copy(porcelain_file("examples/add2"), tmp, recursive = TRUE)
path <- file.path(tmp, "add2")
```


```{r, echo = FALSE, comment = NA}
withr::with_dir(dirname(path), fs::dir_tree("add2"))
```

As for the simple example in `vignette("porcelain")` we have written the api into a single file `R/api.R` but this could be split over as many files as you prefer

```{r, echo = FALSE, results = "asis"}
show_file(file.path(path, "R/api.R"), "r")
```

Our `DESCRIPTION` file includes the roxygen2 setup

```{r, echo = FALSE, results = "asis"}
show_file(file.path(path, "DESCRIPTION"), "")
```

We include a small json schema `inst/schema/numeric.json`:

```{r, echo = FALSE, results = "asis"}
show_file(file.path(path, "inst/schema/numeric.json"), "json")
```

Running `roxygen2::roxygenize(path)` on the file will build the interface

```{r}
roxygen2::roxygenize(path)
```

The contents now include `R/porcelain.R` (and `man`, created by roxygen2 but empty)

```{r, echo = FALSE, comment = NA}
withr::with_dir(dirname(path), fs::dir_tree("add2"))
```

The `R/porcelain.R` file contains automatically-generated endpoint definitions

```{r, echo = FALSE, results = "asis"}
show_file(file.path(path, "R/porcelain.R"), "r")
```
